---
title: "hw5"
output: html_document
---

```{r}
setwd("C:\\Users\\Matth\\Desktop\\Economic Forecasting\\R\\Week 10")
rm(list=ls(all=TRUE))

library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
library(nlstools)
library(fpp)
library(strucchange)
library(Quandl)
library(zoo)
library(PerformanceAnalytics)
library(quantmod)
library(vars)
library(lmtest)
library(dlnm)
library(hts)
library(tseries)
library(rugarch)
library(TSA)
library(fGarch)
library(pastecs)
```

```{r}


#Setting up low persistence simulation
spec <- ugarchspec(
  variance.model = list(model = "sGARCH",
                        garchOrder = c(1,1)),
  mean.model = list(armaOrder = c(1,1), include.mean = TRUE),
  distribution.model = "norm",
  fixed.pars = c(mu = 2, ma1 = .2, ar1 = .3, omega = 2, alpha1 = .4, beta1 = .4))
lowpersistence <- ugarchpath(spec, n.sim = 1000, n.start = 1)

#Plotting time series
plot(fitted(lowpersistence), type = "l")
#Plotting conditional variance and conditional standard deviation
sigma2_low <- (sigma(lowpersistence))^2
plot(sigma(lowpersistence), type = "l")
plot(sigma2_low, type = "l")

#histogram of original time series and standardized time series
hist(fitted(lowpersistence))
#autocorrelograms of original time series, square of the series, standardized series, and square of the standardized series



#Setting up high persistence simulation
spec2 <- ugarchspec(
  variance.model = list(model = "sGARCH",
                        garchOrder = c(1,1)),
  mean.model = list(armaOrder = c(0,1), include.mean = TRUE),
  distribution.model = "norm",
  fixed.pars = c(mu = 2, ma1 = .3, omega = 2, alpha1 = .1, beta1 = .88))
highpersistence <- ugarchpath(spec2, n.sim = 1000, n.start = 1)

#plotting time series
plot(fitted(highpersistence), type = "l")

#plotting conditional variance and standard deviation
sigma2_high <- (sigma(highpersistence))^2
plot(sigma(highpersistence), type = "l")
plot(sigma2_high)

#histogram of original time series and standardized time series
hist(fitted(highpersistence))

#autocorrelograms of original time series, square of the series, standardized series, and square of the standardized series

```

### Number 2, 14.3

```{r}
rm(list=ls(all=T))
df <- read.csv("14_3data.csv")


returns <- as.data.frame(diff(log(df$Adj.Close)))

returns$returns2 <- (returns$`diff(log(df$Adj.Close))`)^2
plot(returns$`diff(log(df$Adj.Close))`, type = "l")
#Volatility was initially somewhat high, then volatility dropped to very low levels and remained constant for a while. Volatility then picked up and peaked at extremely high levels, and then went back down to medium levels where it remains.
acf(returns$`diff(log(df$Adj.Close))`)
pacf(returns$`diff(log(df$Adj.Close))`)
acf(returns$returns2)
pacf(returns$returns2)

archmodel = ugarchspec(
  variance.model = list(model ="sGARCH", garchOrder = c(9,0)),
  mean.model = list(armaOrder = c(2,2), include.mean = TRUE),
  distribution.model = "sstd")

archmodel@model$pars

modelfit=ugarchfit(spec=archmodel, data = returns$`diff(log(df$Adj.Close))`)

plot(modelfit@fit[["residuals"]], type = "l")

acf(modelfit@fit[["residuals"]])
pacf(modelfit@fit[["residuals"]])


```

### Problem 3 14.4

```{r}
?ugarchforecast

forecast <- ugarchforecast(modelfit, data = NULL, n.ahead = 2, n.roll = 0, out.sample = 0)
#need to add confidence intervals 
```


(Removing conditional mean dynamics before modeling volatility dynamics) In
the application in the text, we noted that NYSE stock returns appeared to have
some weak conditional mean dynamics, yet we ignored them and proceeded
directly to model volatility.
a. Instead, first fit autoregressive models using the SIC to guide order selection,
and then fit GARCH models to the residuals. Redo the entire empirical
analysis reported in the text in this w^y, and discuss anv important differences
in the results.
b. Consider instead the simultaneous estimation of all parameters of AR(p)-
GARCH models. That is, estimate regression models where the regressors are
lagged dependent variables and die disturbances display GARCH. Redo the
entire empirical analysis reported in the text in this way, and discuss any
important differences in the results relative to those in the text and those
obtained in part a. 

### Problem 4 14.3 diebold

```{r}
rm(list=ls(all=TRUE))
getSymbols("^NYA", from = "1988-01-01", to = "2001-12-31")

adj.close <- NYA$NYA.Adjusted


returns <- diff(log(adj.close))

returns <- returns[-1,]
returns2 <- returns^2
returnsts <- as.ts(returns)
returns2ts <- as.ts(returns2)
returnszoo <- as.zoo(returns)
returns2zoo <- returnszoo^2


#plotting
plot(returnszoo, type = "l")
plot(returns2zoo, type = "l")

hist(returnszoo)

acf(returnsts, lag.max = 1000)
pacf(returnsts, lag.max = 1000)
auto.arima(returnszoo)



m1 <- arima(returnszoo, order = c(1,0,0))
m2 <- arima(returnsts, order = c(3,0,0))
m3 <- arima(returnszoo, order = c(5,0,5))

acf(m2$residuals, lag.max = 300)
pacf(m2$residuals, lag.max = 300)

acf(returns2ts)
pacf(returns2ts)

m4 <- auto.arima(returnsts)

###continuing...

m4residuals <- m4$residuals

plot(m4residuals, type = "l")
acf(m4residuals)
pacf(m4residuals)
#starting garch model

model=ugarchspec(
variance.model = list(model = "sGARCH", garchOrder = c(2, 2)),
mean.model = list(armaOrder = c(5, 2), include.mean = TRUE),
distribution.model = "sstd")

modelfit=ugarchfit(spec=model,data=returnszoo)
modelfit
plot(modelfit)





```





4. (Direct modeling of volatility proxies) In the text, we fit an AR(5) directly to a
subset of the squared NYSE stock returns. In this exercise, use the entire KYSE
dataset.
a. Construct, display, and discuss the fitted volatility series from the AR(5)
model.
b. Construct, display, and discuss an alternative fitted volatility series obtained by
exponendal smoothing, using a smoothing parameter of 0.10, corresponding
to a large amount of smoodting, but less than that done in the text.
c. Construct, display, and discuss the volatility series obtained by fitting an
appropriate GARCH model.
d. Contrast the results of parts a, b, and c.
e. Why is fitting of a GARCH model preferable in principle to die AR(5) or
exponential smoothing approaches? 

###Problem 4, 14.4 Diebold

```{r}


```

5. (GARCH volatility forecasting) You work for Xanadu, a luxury resort in the
tropics. The daily temperature in the region is beautiful year-round, with a mean
around 76 (Fahrenheit!) and no conditional mean dynamics. Occasional pressure
systems, however, can cause bursts of temperature volatility. Such volatility bursts
generally don't last long enough to drive away guests, but the resort still loses
revenue from fees on activities that are less popular when the weather isn't
perfect. In the middle of such a period of high temperature volatility, your boss
gets worried and asks vou make a forecast of volatility over the next 10 days. After
some experimentation, you find that daily temperature y , follows
y, | f2,_i ~ AT (p.. of) ,
>>
where a," follows a GARCH(1, 1) process,
t 2 " 2
a , = w 4- aE,_, + ficr,_l .
a. Estimation of your model using historical daily temperature data yields
ji = 76, u> = 3, a = 0.6, and |3 = 0. If yesterday's temperature was 92 degrees,
generate point forecasts for each of the next 10 days' conditional variance.
b. According to your volatility forecasts, how many days will it take until volatility
drops enough such that there is at least a 90% probability that the
temperature will be within 4 degrees of 76?
c. Your boss is impressed by vour knowledge of forecasting and asks you whether
your model can predict the ne



###Problem 5, 14.5 Diebold


```{r}

specweath <- ugarchspec(
  variance.model = list(model = "sGARCH",
                        garchOrder = c(1,1)),
  mean.model = list(armaOrder = c(0,0), include.mean = TRUE),
  distribution.model = "norm",
  fixed.pars = c(mu = 76, omega = 3, alpha1 = .6, beta1 = 0))
weather <- ugarchpath(specweath, n.sim = 1000, n.start = 1)

plot(fitted(weather), type = "l")

#Still need to figure out how to forecast with this object type, also need to figure out how to condition forecast off previous observation being 96 degrees





```